---
title: "Introduction"
author: Pascal Schmidt
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, message = FALSE, warning = FALSE, echo=FALSE}
knitr::opts_chunk$set(
  message = F,
  warning = F,
  comment = "#>",
  fig.height = 6, 
  fig.width = 8
)

library(dplyr)
library(SveltePlots)
library(ggplot2)
```

# Introduction

- The main function in `SveltePlots` is the `sp()` function where we can specify the chart type, `line`, `points`/`scatter`, `bar`, `hist`, and `density`.
- All functions start with `sp_`.
- Series can be drawn individually with `sp_add_series()` and it currently supports `points`, `line`, or `bands`.
- There is also a `spaes()` function which works similar to the `aes()` function in `ggplot2`.

# Scatter Plots

The first chart we are building is a scatter plot. We are providing the data frame, the mapping, and the type of the plot and the `sp()` function will plot the chart. 

```{r}
#| column: screen-inset
data("penguins")
sp(
  data = penguins,
  mapping = spaes(x = flipper_length_mm, y = bill_length_mm, group = species),
  type = "points"
) %>% sp_title("Penguins")
```

We can also specify different colors, the size of the points, the opacity, if we want tooltips and a legend, and the height of the chart. 

```{r}
sp(
  data = penguins,
  mapping = spaes(x = flipper_length_mm, y = bill_length_mm, group = species),
  type = "points",
  size = 3, 
  alpha = 1,
  colors = c("red", "blue", "green"),
  height = 500
)
```

```{r}
n <- 500  

variable1 <- rnorm(n, mean = 0, sd = 1)  
variable2 <- rnorm(n, mean = 5, sd = 2) 

big_data <- data.frame(Variable1 = variable1, Variable2 = variable2)

sp(
  data = big_data,
  type = "points",
  mapping = spaes(x = Variable1, y = Variable2), 
  alpha = 0.5,
  height = 500,
  include_legend = FALSE
)
```


# Line Charts

```{r}
data("walmart_sales_weekly")
walmart_sales_weekly <- walmart_sales_weekly %>%
  dplyr::select(Dept, Date, Weekly_Sales) %>%
  dplyr::arrange(Date) %>%
  dplyr::as_tibble() %>%
  dplyr::mutate(
    Date = lubridate::ymd(Date)
  )

sp(
  data = walmart_sales_weekly,
  type = "line",
  mapping = spaes(x = Date, y = Weekly_Sales, group = Dept)
) %>% 
  sp_y_axis(format = "$,.3r") %>% 
  sp_title("Walmart Weekly Sales")
```

Individual series can be added with the `sp_add_series()` function. For individual series, we can also specify if we want the legend and also tooltip for the series. By default, the plots displays the tooltip and also legend.

```{r}
data("purchases")
sp(
  data = purchases_age, 
  type = "points",
  mapping = spaes(x = adjusted_timestamp, y = revenue_roll, group = age)
) %>%
  sp_add_series(
    data = purchases_age[purchases_age$revenue == max(purchases_age$revenue), ], 
    mapping = spaes(x = adjusted_timestamp, y = revenue, group = age),
    type = "points",
    size = 5
  ) 
```

# Bar Charts

```{r}
data("fruit")
sp(
  data = fruit[, c("year", "fruit", "value")] %>% 
    dplyr::filter(fruit != "bananas"),
  mapping = spaes(x = year, y = value, group = fruit),
  type = "bar",
  mode = "grouped",
  height = 500
)  %>% 
  sp_add_series(
    data = fruit[, c("year", "fruit", "value")] %>% 
      dplyr::filter(fruit == "bananas"),
    mapping = spaes(x = year, y = value, group = fruit),
    type = "bar",
  )

sp <- sp(
  data = fruit[, c("year", "fruit", "value")],
  mapping = spaes(x = year, y = value, group = fruit),
  type = "bar",
  mode = "stacked", height = 500
) %>% 
  sp_tooltip(
    type = "vertical_line"
  )
sp

sp <- sp(
  data = fruit[, c("year", "fruit", "value")],
  mapping = spaes(x = year, y = value, group = fruit),
  type = "bar",
  mode = "percent", height = 500
) %>% 
  sp_tooltip(
    type = "vertical_line"
  )
sp

sp(
  data = mtcars %>% tibble::rownames_to_column() %>% dplyr::slice(1:10),
  mapping = spaes(x = rowname, y = mpg),
  type = "bar", include_legend = FALSE
) %>%  
  sp_x_axis(
    label = "Date", rotation_axis_ticks = -30,
    text_anchor = "start"
  )
```

# Density Charts

```{r}
data("gapminder")

gapminder <- gapminder %>% 
  dplyr::mutate(
    country = as.character(country),
    year = lubridate::ymd(paste0(year, "-01-01"))
  )

sp(
  data = gapminder %>% 
    dplyr::filter(
      year == lubridate::ymd("2007-01-01"), 
      continent != "Oceania"
    ),
  mapping = spaes(x = lifeExp, group = continent),
  type = "density", alpha = 1, size = 1
)

sp(
  data = mtcars,
  mapping = spaes(x = mpg),
  type = "density", alpha = 0.5, size = 1
)
```

# Histograms

```{r}
sp(
  data = gapminder %>% dplyr::filter(year == lubridate::ymd("2007-01-01"), continent != "Oceania"),
  mapping = spaes(x = lifeExp, group = continent),
  type = "histogram", alpha = 0.5, size = 1
)

sp(
  data = mtcars,
  mapping = spaes(x = mpg),
  type = "histogram", alpha = 0.5, size = 1
) 

sp(
  data = diamonds,
  mapping = spaes(x = price),
  type = "histogram", alpha = 0.5, size = 1, breaks = 50
)
```





